Epistemic Semantics in Guarded String Models

### 2020-09-22 tex/sdraft.tex
center & shrink figures
tighten figures

\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
\input{syntax.tex}
\usepackage{wasysym,graphicx}

@@

\begin{figure}
  \(\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ b }^{\epikspec} && \mathcal A^{b\varphi}_{\mathsf B}\\
  \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
  \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
  \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
  \sem{p^*}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
  \sem{ \neg p }^{\epikspec} && \sem{\bullet^*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
  \end{array}
  \)

@@

\begin{figure}
\begin{hangingpar}
{\tt St} \hspace{2mm} Tests such as 0 1 1 0. The length is the number of generators.
\end{hangingpar}

@@

\begin{figure} \label{relclosure}
{\raggedright
\begin{hangingpar}
{\tt{define RelKpl(R)}}
$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
$\underbrace{\mbox{\tt{[R+]}}}_a$
$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
\end{hangingpar}}

@@

\begin{hangingpar}
{\tt{define Kst(R)
    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}} The Fst
operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
{\tt{R}}.
\end{hangingpar}

@@

We also implement the semantics of Epik terms using lazy
lists~\cite{LazyLists} in Haskell, rather than the direct
interpretation as sets. Unfortunately, regular expressions, and hence
also Epik, can denote infinite sets of strings, for example the term
$\_^*$. Normally regular languages are represented using a finite
coalgebra~\cite{}. However the non-distributivity of $\diamond$
accross $;$ complicates this construction\footnote{Axiomatic and
  Coalgebraic models for Epik are open questions}. To sidestep this,
we parameterize the interpretation function on a positive integer $n$
and only produce guarded strings of length $n$ or less.

We translate the Epik terms into a Haskell algebraic datatype that
represents terms with the same signature as described in Section 2. We
then paramaterize the interpretation function, $\llparenthesis p
\rrparenthesis^{\varphi,{\mathcal{E}}}_n$ on an integer $n$ that
describes the maximum length of a string we will produce.

@@

We lift the remaining operators (except Kleene star) to their list equivalents:
union becomes list append (written $\texttt{+\!+}$); concatenation becomes the
fusion product (lifted to lists this time), negation is implemented using list
difference ($\setminus$), and the modal operator lifts the alternative relation
over lists of strings\footnote{Figure~\ref{fig:list-interp} depicts this using
  the list comprehension notation, which is analogous to set builder notation,
  except that it is written using square brackets. The order of elements is
  preserved indicated by the keyword \texttt{for}}. The only caveat to these
direct interpretations is that we restrict the operators to have size $\leq n$,
denoted as $l|_n$ for a list of guarded strings $l$.

@@

\begin{figure}
  \[\begin{array}{r >{\triangleq}c l}
  \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_0 && [] \\
  \llparenthesis 0 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] \\
  \llparenthesis 1 \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^\varphi \rfloor \\
  \llparenthesis e \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \hat{\mathcal E}^\varphi(e)\rfloor \\
  \llparenthesis \psi \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \lfloor \mathcal A_{\mathsf B}^{\varphi~\psi} \rfloor \\
  \llparenthesis p + q \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \mathbin{\texttt{+\!+}} \llparenthesis q \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
  \llparenthesis p;q \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  \left(\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond \llparenthesis q\rrparenthesis^{\varphi,{\mathcal{E}}}_n\right)|_n\\
  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n && [] + \left(\llparenthesis p\rrparenthesis^{\varphi,{\mathcal{E}}}_n \diamond  \llparenthesis p^* \rrparenthesis^{\varphi,{\mathcal{E}}}_{m}\right)|_n \\
  \multicolumn 3 r {\text{where}~
    \begin{array}[t]{r >{\hspace{-.7em}}l}
      i &= \max\{1,\min\{|g|\mid g \in \llparenthesis p \rrparenthesis_n^{\varphi,{\mathcal{E}}}\}\}\\
      m &= \max\{0, n - i\}
    \end{array}}\\
  \llparenthesis \neg p \rrparenthesis^{\varphi,{\mathcal{E}}}_n && \llparenthesis \_^* \rrparenthesis^{\varphi,{\mathcal{E}}}_n \setminus \llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n \\
  \llparenthesis \Diamond_a~p \rrparenthesis^{\varphi,{\mathcal{E}}}_n &&
  [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\llparenthesis p \rrparenthesis^{\varphi,{\mathcal{E}}}_n]
  \end{array}
  \]
  \caption{Bounded interpretation using lazy lists}
  \label{fig:bound-interp}
\end{figure}

@@

>>>>

\newcommand{\dom}{\ensuremath{\mathrm{dom}}}
\input{syntax.tex}
\usepackage{wasysym,graphicx}
\usepackage{tcolorbox}

@@

\begin{figure}
  \centering
  \(\begin{array}{l >{\triangleq}c l}
  \sem{ 0 }^{\epikspec} && \emptyset \\  
  \sem{ 1 }^{\epikspec} && \mathcal A^\varphi_{\mathsf B}\\
  \sem{ b }^{\epikspec} && \mathcal A^{b\varphi}_{\mathsf B}\\
  \sem{ \bar{\sigma} }^{\epikspec} && \AphiB \setminus \sem{\sigma}^{\epikspec}\\
  \sem{ e }^{\epikspec}&&  \setabs{\alpha e \beta}{\alpha \zeta_e \beta} \\
  \sem{{p+q}}^{\epikspec} && \sem{{p}}^{\epikspec} \cup \sem{ q }^{\epikspec}\\
  \sem{{pq}}^{\epikspec} && \setabs{x \diamond y}{ \begin{array}{l}x \in \sem{p}^{\epikspec} \\ y \in \sem{q}^{\epikspec}\\\mbox{$x \diamond y$ is defined}\end{array}} \\
  \sem{p^*}^{\epikspec} && \displaystyle\bigcup_{n \geq 0} {\sem{ p^n }^{\epikspec}} \\
  \sem{ \neg p }^{\epikspec} && \sem{\bullet^*}^{\epikspec} \setminus \sem{ p }^{\epikspec} \\
  \end{array}
  \)

@@

\begin{figure}
\footnotesize
\begin{hangingpar}
{\tt St} \hspace{2mm} Tests such as 0 1 1 0. The length is the number of generators.
\end{hangingpar}

@@

\begin{figure} \label{relclosure}
  \footnotesize
{\raggedright
  \begin{hangingpar}
{\tt{define RelKpl(R)}}\\
$\underbrace{\mbox{\tt{Squash.i\hspace{1pt}.o.}}}_c$
$\underbrace{\mbox{\tt Wf0\hspace{1pt}.o.}}_b$
$\underbrace{\mbox{\tt{[R+]}}}_a$
$\underbrace{\mbox{\tt .o.\hspace{1pt}Wf0}}_b$
$\underbrace{\mbox{\tt{.o.\hspace{1pt}Squash}}}_c$
\end{hangingpar}}

@@

\begin{hangingpar}
{\tt{define Kst(R)
    [St\hspace{1pt}.x.\hspace{1pt}St]\hspace{1pt}|\hspace{1pt}Kpl(X);}}\\ The Fst
operation {\tt{.x.}} is Cartesian product.  {\tt{R.i}} is the inverse of relation
{\tt{R}}.
\end{hangingpar}

@@

We also implement the semantics of Epik terms using lazy
lists~\cite{LazyLists} in Haskell, rather than the direct
interpretation as sets. Unfortunately, regular expressions, and hence
also Epik, can denote infinite sets of strings, for example the term
$\bullet^*$. Normally regular languages are represented using a finite
coalgebra~\cite{}. However the non-distributivity of $\diamond$
accross $;$ complicates this construction\footnote{Axiomatic and
  Coalgebraic models for Epik are open questions}. To sidestep this,
we parameterize the interpretation function on a positive integer $n$
and only produce guarded strings of length $n$ or less.

We translate the Epik terms into a Haskell algebraic datatype that
represents terms with the same signature as described in Section 2. We
then paramaterize the interpretation function on an integer $n$ that
describes the maximum length of a string we will produce.

@@

We lift the remaining operators (except Kleene star) to their list
equivalents: union becomes list append (written $\texttt{+\!+}$);
concatenation becomes the fusion product (lifted to lists this time),
negation is implemented using list difference ($\setminus$), and the
modal operator lifts the alternative relation over lists of
strings\footnote{Figure~\ref{fig:list-interp} depicts this using the
  list comprehension notation, which is analogous to set builder
  notation, except that it is written using square brackets. Element
  order is evoked by the keyword \texttt{for}, rather than using the
  unordered $\forall$.}. The only caveat to these direct
interpretations is that we restrict the operators to have size $\leq
n$, denoted as $l|_n$ for a list of guarded strings $l$.

@@

\newcommand{\semlst}[2]{\llparenthesis#1\rrparenthesis^{\mathsf B, \mathsf P, \phi, \zeta}_{#2}}
\newcommand{\setToList}[1]{\lfloor#1\rfloor}

\begin{figure}
  \centering
  \(\begin{array}{r >{\triangleq}c l}
  \semlst p 0 && [] \\
  \semlst 0 n && [] \\
  \semlst 1 n && \setToList \AphiB \\
  \semlst e n && [ \alpha e \beta \mid \alpha \mathrel{\zeta_e} \beta ] \\
  \semlst b n && \setToList{ \mathcal A_{\mathsf B}^{b\psi}} \\
  \semlst{p + q} n && \semlst p n \mathbin{\texttt{+\!+}} \semlst q n \\
  \semlst{p;q} n &&
  \left(\semlst p n \diamond \semlst q n \right)|_n\\
  \semlst{p^*} n && [] + \left(\semlst p n \diamond \semlst{p^*} {n-i}\right) |_n \\
  \multicolumn 3 r {\text{where}~
    \begin{array}[t]{r >{\hspace{-.7em}}l}
      i &= \max\{1,\min\{|g|\mid g \in \semlst p n \}\} \\
    \end{array}}\\
  \semlst{\neg p} n && \semlst {\bullet^*} n \setminus \semlst p n \\
  \semlst{\Diamond_a~p } n &&
         [g' \mid g'~\hat{R_a}~g,~\texttt{for}~g~\texttt{in}~\semlst p n]
  \end{array}
  \)
  \caption{Bounded interpretation using lazy lists}
  \label{fig:bound-interp}
\end{figure}

@@

\section{Conclusion}

\ehc{Insert happy words!}
